---
title: "Relaxed Lasso multinomial LR simulation with multinomial data"
author: "Alexander Romanus"
knit: (function(inputFile, encoding) { 
      out_dir <- 'development_results/final_implementation_knits';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, paste(runif(1), 'relaxed_simulation_multinomial.html'))) })
output:
  html_document: 
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
library(casebase)
library(future.apply)
library(glmnet)
library(parallelly)
library(timereg)
library(mtool)
library(parallel)
library(tictoc)
library(tidyverse)
library(cmprsk)
library(survsim)
library(caret)
library(Matrix)
library(dplyr)
library(ggplot2)
library(grid)
library(knitr)

source("../../../src/final_relaxed_implementation.R")
source("../../../src/fitting_functions.R")
```

#Check cv.glmnet(relax=TRUE) vs mtool relaxed implementation on multinomial data

### Setup n > p (n = 400, p = 120)
``` {r setup-simulation,echo=FALSE}
p = 120
n = 400
nfolds = 5
seed = 2023
set.seed(seed)

# Generate covariates 
X <- matrix(rnorm(n * p), n, p)

# coefficients for each choice
beta0 <- rep(0, p)
beta1 <- c(rep(3, p/2), rep(0, p/2))
zero_beta1 <- which(beta1 == 0)

beta2 <- c(rep(3, p/2), rep(0, p/2))
zero_beta2 <- which(beta2 == 0)


# vector of probabilities
vProb = cbind(exp(X%*%beta0), exp(X%*%beta1), exp(X%*%beta2))

# multinomial draws
mChoices <- t(apply(vProb, 1, rmultinom, n = 1, size = 1))
dfM <- cbind.data.frame(y = apply(mChoices, 1, function(x) which(x == 1)), X)
# Rename covariates 
colnames(dfM)[2:(p+1)] <- paste0('x', 1:p)

train_index_y = caret::createDataPartition(dfM$y, p = 0.80, list = FALSE)

dfM_train = dfM[train_index_y, ]
dfM_test = dfM[-train_index_y, ]

x_train = as.matrix(dfM_train[, 2:(p+1)])
y_train = dfM_train$y - 1

x_test = as.matrix(dfM_test[, 2:(p+1)])
y_test = dfM_test$y - 1

dfM_train_list = list("covariates" = as.matrix(cbind(dfM_train[, c(2:ncol(dfM_train))])),
                     "y" = dfM_train$y - 1)

dfM_test_list = list("covariates" = as.matrix(cbind(dfM_test[, c(2:ncol(dfM_test))])),
                     "y" = dfM_test$y - 1)
```


## glmnet-implementation with no additional penalization (gamma = 0)
``` {r glmnet-implementation, echo=FALSE, results = 'hide'}
alpha <- 1

tryCatch({
  # glmnet relax = TRUE
  fit.glmnet.relaxed <- glmnet::cv.glmnet(
    x = x_train, y = y_train,
    family = "multinomial",
    intercept = TRUE,
    type.multinomial = "grouped",  # same sparsity pattern for all outcome classes
    alpha = 1, nfolds = 10,
    relax = TRUE, gamma = 0)
  
    glmnet_lambda_max = fit.glmnet.relaxed$lambda[1]
    glmnet_lambda_min = fit.glmnet.relaxed$relaxed$lambda.min
    
    fit_val_min_glmnet <- glmnet(x = x_test, y = y_test,
                                 lambda = glmnet_lambda_min,
                                 family = "multinomial",
                                 intercept = TRUE,
                                 type.multinomial = "grouped")
  
  
  glmnet_min1 <- coef(fit_val_min_glmnet)$`1` - coef(fit_val_min_glmnet)$`0`
  glmnet_min2 <- coef(fit_val_min_glmnet)$`2` - coef(fit_val_min_glmnet)$`0`
  
  res_glmnet_min1 <- varsel_perc(glmnet_min1[-1], beta1)
  res_glmnet_min2 <- varsel_perc(glmnet_min2[-1], beta2)
  
  reformat_glmnet_min1 = c(glmnet_min1[-1], glmnet_min1[1])
  reformat_glmnet_min2 = c(glmnet_min2[-1], glmnet_min2[1])
  glmnet_coefs = list("coefficients" = as.matrix(cbind(reformat_glmnet_min1, reformat_glmnet_min2)))
  
  # Test multi-deviance
  dev_glmnet <- multi_deviance_multinomial(data = dfM_test_list, fit_object = glmnet_coefs)
  
  plot(fit.glmnet.relaxed)
    
}, 
error = function(e) {
  print(e)
  }
)

```

## mtool multinomial model fit with Relaxed Lasso with no additional penalization (gamma = 0)
``` {r relaxed-lasso-implementation-no-gamma, echo=FALSE, message=FALSE, results = 'hide'}
fit.mtool.relaxed_no_penalization <- multinomial_casebase_relaxed_lasso(train = dfM_train, gamma = 0, seed = seed)

# Refit on lambda.min
output <- capture.output(fit_val_min_no_pen <- fit_model(dfM_test_list, regularization = 'l1',
                           lambda = fit.mtool.relaxed_no_penalization$lambda.min, alpha = 1, unpen_cov = 1))

coef_val1 <- fit_val_min_no_pen$coefficients[1:p, 1]
coef_val2 <- fit_val_min_no_pen$coefficients[1:p, 2]

# Res
res_mtool_relaxed_min1_no_pen <- varsel_perc(coef_val1, beta1)
res_mtool_relaxed_min2_no_pen <- varsel_perc(coef_val2, beta2)

# Test multi-deviance
dev_mtool_relaxed_no_pen <- multi_deviance_final(data = dfM_test_list, fit_object = fit_val_min_no_pen)

plot_post_relaxed(fit.mtool.relaxed_no_penalization)
```


## mtool multinomial model fit with Relaxed Lasso with small additional penalization (gamma = 0.00001)
``` {r relaxed-lasso-implementation-small-gamma, echo=FALSE, message=FALSE, results = 'hide'}
fit.mtool.relaxed_small_penalization <- multinomial_casebase_relaxed_lasso(train = dfM_train, gamma = 0.00001, seed = seed)

# Refit on lambda.min
output <- capture.output(fit_val_min_small_pen <- fit_model(dfM_test_list, regularization = 'l1',
                               lambda = fit.mtool.relaxed_small_penalization$lambda.min, alpha = 1, unpen_cov = 1))

coef_val1 <- fit_val_min_small_pen$coefficients[1:p, 1]
coef_val2 <- fit_val_min_small_pen$coefficients[1:p, 2]

# Res
res_mtool_relaxed_min1_small_pen <- varsel_perc(coef_val1, beta1)
res_mtool_relaxed_min2_small_pen <- varsel_perc(coef_val2, beta2)


# Test multi-deviance
dev_mtool_relaxed_small_pen <- multi_deviance_final(data = dfM_test_list, fit_object = fit_val_min_small_pen)

plot_post_relaxed(fit.mtool.relaxed_small_penalization)
```



## mtool multinomial model fit with Relaxed Lasso with significant additional penalization (gamma = 0.001)
``` {r relaxed-lasso-implementation-sig-gamma, echo=FALSE, message=FALSE, results = 'hide'}
fit.mtool.relaxed_sig_penalization <- multinomial_casebase_relaxed_lasso(train = dfM_train, gamma = 0.001, seed = seed)

# Refit on lambda.min
output <- capture.output(fit_val_min_sig_pen <- fit_model(dfM_test_list, regularization = 'l1',
                               lambda = fit.mtool.relaxed_sig_penalization$lambda.min, alpha = 1, unpen_cov = 1))

coef_val1 <- fit_val_min_sig_pen$coefficients[1:p, 1]
coef_val2 <- fit_val_min_sig_pen$coefficients[1:p, 2]

# Res
res_mtool_relaxed_min1_significant_pen <- varsel_perc(coef_val1, beta1)
res_mtool_relaxed_min2_significant_pen <- varsel_perc(coef_val2, beta2)


# Test multi-deviance
dev_mtool_relaxed_significant_pen <- multi_deviance_final(data = dfM_test_list, fit_object = fit_val_min_sig_pen)

plot_post_relaxed(fit.mtool.relaxed_sig_penalization)
```

## Final results
``` {r results, echo=FALSE, message=FALSE}
Res <- rbind(res_mtool_relaxed_min1_no_pen, res_mtool_relaxed_min2_no_pen,
             res_mtool_relaxed_min1_small_pen, res_mtool_relaxed_min2_small_pen, res_mtool_relaxed_min1_significant_pen,
             res_mtool_relaxed_min2_significant_pen)

devs <- rbind(dev_mtool_relaxed_no_pen, dev_mtool_relaxed_small_pen,
            dev_mtool_relaxed_significant_pen)
  

# If glmnet model was fit, include its coefficient estimates and deviances in final results tables
if(exists("res_glmnet_min1")) {
  Res <- rbind(res_glmnet_min1, res_glmnet_min2, Res)
  devs <- rbind(dev_glmnet, devs)
    
  rownames(Res) <- c("glmnet.lambda.min_reference1", "glmnet.lambda.min_reference2", "mtool.relaxed.no.pen.lambda.min_reference1",
                     "mtool.relaxed.no.pen.lambda.min_reference2", "mtool.relaxed.small.pen.lambda.min_reference1",
                     "mtool.relaxed.small.pen.lambda.min_reference2", "mtool.relaxed.significant.pen.lambda.min_reference1",
                     "mtool.relaxed.significant.pen.lambda.min_reference2")
  
  
  rownames(devs) <- c("glmnet.lambda.min",  "mtool.relaxed.no.pen.lambda.min",
                     "mtool.relaxed.small.pen.lambda.min",
                     "mtool.relaxed.significant.pen.lambda.min")
} else {
  rownames(Res) <- c("mtool.relaxed.no.pen.lambda.min_reference1",
                     "mtool.relaxed.no.pen.lambda.min_reference2", "mtool.relaxed.small.pen.lambda.min_reference1",
                     "mtool.relaxed.small.pen.lambda.min_reference2", "mtool.relaxed.significant.pen.lambda.min_reference1",
                     "mtool.relaxed.significant.pen.lambda.min_reference2")
  
  
  rownames(devs) <- c("mtool.relaxed.no.pen.lambda.min",
                     "mtool.relaxed.small.pen.lambda.min",
                     "mtool.relaxed.significant.pen.lambda.min")
}

kable(Res)
kable(devs)
```