%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

First and foremost, I would like to acknowledge my supervisor, Prof. Gabriela V. Cohen Freue, for her valuable support and insightful feedback throughout my M.Sc. I am also thankful to Prof. Paul Gustafson for taking the time to read and evaluate my thesis. I would also like to thank Dr. Yi Lian for his helpful advice. I am also very thankful to Dr. Sahir Bhatnagar, Dr. Maxime Turgeon and to the whole CANSII CRT group, for the fruitful collaboration on \texttt{casebase}, and for the financial support through my MSc. Additionally, I want to thank Maxwell Douglas, Evan Popoff and Basia Rogula for their support during my co-op work terms. I also wish to extend my sincere gratitude to my undergraduate supervisor Dr. Ian Dworkin, who actively gave me the confidence and encouragement to pursue grad school and research.
\bigskip \par
I am grateful for all of my wonderful fellow classmates and friends both in and outside the Department of Statistics. I want to thank Caitlin, Mansi, and Sangay in particular for the many good times. I also wish to acknowledge my fellow lab members, Chloe, Giuseppe, Maggie, and Jana, in particular for coding and writing along with me everyday in Koerner. I would also like to thank Dr. Anthony Christidis for his helpful advice on my simulations. I am very grateful to the Department of Statistics for their prompt support and help, and for their financial support through the Rick White Award. I also want to thank Alex Romanus for contributing to my project as a part of his USRA.
\bigskip \par
Finally, I am thankful for my family. I would not have been able to accomplish any of this without their love and support.

\bigskip \par


